---
title: "Introduction to Quantstrat"
output: html_notebook
---

Experimentation with `quantstrat` via the https://quantstrattrader.wordpress.com/ 
blog.

```{r setup}
#install.packages("quantstrat", repos="https://r-forge.r-project.org")
#devtools::install_github("ilyakipnis/IKTrading")
#devtools::install_github("ilyakipnis/DSTrading")A

library(quantstrat)
library(IKTrading)
#library(tidyverse)

options("getSymbols.auto.assign" = FALSE)
options("getSymbols.warning4.0" = FALSE)

```

# Getting and Working with Data

We will work on the S&P 500 index.

```{r get_stock_data}
SPY <- getSymbols("SPY", from = "2007-01-01", to = "2017-01-01")
#glimpse(dat)
head(SPY)
tail(SPY)
```
```{r}
chart_Series(SPY)
```

`zoom_Chart` does work inside of notebooks, but we'll also use standard 
`xts` subsetting in a few places for demo purposes.
```{r zoom_2009}
zoom_Chart("2009")
#chart_Series(SPY["2009"])
```
Add Bolinger Bands
```{r}
add_BBands()
```

Zooming into a specific month
```{r}
chart_Series(SPY["2015-03"])
```

Calculate 200 day simple moving average (SMA, via the `TTR` package, a 
dependency of the `quantstrat` package) on the close `Cl` price.

```{r calculate_sma}
sma <- SMA(Cl(SPY), n = 200)
tail(sma)
```

Add the SMA to the chart.
```{r}
plot(add_TA(sma, on = 1))
```

Calculate the RSI
```{r calculate_rsi}
rsi <- RSI(price = Cl(SPY), n = 2)
tail(rsi)
```

Add the RSI to the chart as well
```{r}
add_TA(rsi, on=1)
```

# Strategy Exploration

Establish some basic parameters for our strategy.

```{r strategy_parameters}
currency('USD')  
Sys.setenv(TZ = 'UTC')
symbols <- "SPY"
stock(symbols, currency = "USD", multiplier = 1)
init_date <- "1990-01-01" # need to define a date that occurs before the earlist point in our data set
tradeSize <- 10000
initEq <- tradeSize*length(symbols)

```

Initialize the strategy placeholders.

```{r strategy_boilerplate}
strategy.st <- portfolio.st <- account.st <- "RSI_10_6"
rm.strat(portfolio.st)
rm.strat(strategy.st)
# init portfolio
initPortf(portfolio.st, symbols = symbols, initDate = init_date, 
          currency = 'USD', initEq = initEq)
# init accounts
initAcct(account.st, portfolios = portfolio.st, initDate = init_date, 
         currency = 'USD', initEq = initEq)
# init orders
initOrders(portfolio.st, initDate = init_date)
# init strategy
strategy(strategy.st, store = TRUE)
```

Parameter setup
```{r parameters}
#RSA params
nRSI <- 2
thresh1 <- 10
thresh2 <- 6

# SMA parameters
nSMAexit = 5
nSMAfilter <- 200

#order sizing function (IKTrading::osDollarATR)
period <- 10 # days
pctATR <- 0.02 # notional trade size relative to risk
maxPct <- 0.04 # 
```

## Setup Indicators
```{r}
#lagATR is from the IKTrading library
#HLC gives the High-Low-Close
add.indicator(strategy.st, name = "lagATR", 
              arguments = list(HLC = quote(HLC(mktdata)), n = period),
              label = "atrX")
add.indicator(strategy.st, name = "RSI", 
              arguments = list(price = quote(Cl(mktdata)), n = nRSI),
              label="rsi")
add.indicator(strategy.st, name = "SMA", 
              arguments = list(x=quote(Cl(mktdata)), n=nSMAexit),
              label="quickMA")
add.indicator(strategy.st, name = "SMA", 
              arguments = list(x=quote(Cl(mktdata)), n=nSMAfilter),
              label="filterMA")
```
## Setup Signals

```{r}
add.signal(strategy.st, name = "sigComparison",
           arguments = list(columns = c("Close", "filterMA"), relationship = "gt"),
           label = "upTrend")

#sigThreshold compares against a static quantity threshold
add.signal(strategy.st, name = "sigThreshold",
           arguments = list(column = "rsi", threshold = thresh1, relationship = "lt", 
                            cross = FALSE),
           label = "rsiThresh1")
add.signal(strategy.st, name = "sigThreshold",
           arguments = list(column = "rsi", threshold = thresh2, relationship = "lt", 
                            cross = FALSE),
           label = "rsiThresh2")

#sigAND from IKTrading
add.signal(strategy.st, name = "sigAND",
           arguments = list(columns = c("rsiThresh1", "upTrend"), cross = TRUE),
           label = "longEntry1")
add.signal(strategy.st, name = "sigAND",
           arguments = list(columns = c("rsiThresh2", "upTrend"), cross = TRUE),
           label = "longEntry2")

#sigCrossover, unlike sigComparison, only triggers when the relationship goes from FALSE to TRUE
add.signal(strategy.st, name = "sigCrossover",
           arguments = list(columns = c("Close", "quickMA"), relationship = "gt"),
           label = "exitLongNormal")
add.signal(strategy.st, name = "sigCrossover",
           arguments = list(columns = c("Close", "filterMA"), relationship = "lt"),
           label = "exitLongFilter")
```

## Setup Rules

maxPCtATR | maximum position size
osFUN     | order size function (exclusive with `orderqty`)

```{r}

add.rule(strategy.st, name="ruleSignal",
         arguments=list(sigcol="longEntry1",
                        sigval=TRUE,
                        ordertype="market",
                        orderside="long",
                        replace=FALSE,
                        prefer="Open",
                        osFUN=osDollarATR,
                        tradeSize=tradeSize,
                        pctATR=pctATR,
                        maxPctATR=pctATR,
                        atrMod="X"),
         type="enter", path.dep=TRUE, label="enterLong1")
add.rule(strategy.st, name="ruleSignal",
         arguments=list(sigcol="longEntry2",
                        sigval=TRUE,
                        ordertype="market",
                        orderside="long",
                        replace=FALSE,
                        prefer="Open",
                        osFUN=osDollarATR,
                        tradeSize=tradeSize,
                        pctATR=pctATR,
                        maxPctATR=maxPct,
                        atrMod="X"),
         type="enter", path.dep=TRUE, label="enterLong2")

#exit rules
add.rule(strategy.st, name="ruleSignal",
         arguments=list(sigcol="exitLongNormal",
                        sigval=TRUE,
                        orderqty="all",
                        ordertype="market",
                        orderside="long",
                        replace=FALSE,
                        prefer="Open"),
         type="exit", path.dep=TRUE, label="normalExitLong")
add.rule(strategy.st, name="ruleSignal",
         arguments=list(sigcol="exitLongFilter",
                        sigval=TRUE,
                        orderqty="all",
                        ordertype="market",
                        orderside="long",
                        replace=FALSE,
                        prefer="Open"),
         type="exit", path.dep=TRUE, label="filterExitLong")
```
## Run Strategy
```{r}
t1 <- Sys.time()
out <- applyStrategy(strategy=strategy.st, portfolios = portfolio.st, mktdata = SPY)
t2 <- Sys.time()
print(t2 - t1)
```

Set up analytics
```{r}
updatePortf(portfolio.st)
dateRange <- time(getPortfolio(portfolio.st)$summary)[-1]
updateAcct(account.st, dateRange)
updateEndEq(account.st)

```

